### 2019-11-27 Rcode/supplemental_materials.Rmd
editted main text; added goals; added initial simulations

---
title: "Post shock prediction simulations"
author: "Daniel Eck"
date: "November 26, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We load in the R packages that we need.

```{r, message=FALSE}
library(matrixStats)
library(parallel)
library(doParallel)
library(foreach)
library(tidyverse)
```

## Simulations with simple model

```{r, echo = FALSE}
simstudy_normal <- function(n = 20, T = 20, mu.alpha = 2, 
                     sigma.alpha = 2, sigma.X = 1, sigma = 1){

  Tstar <- floor(T/2)
  eta <- rnorm(n+1)
  theta <- rnorm(n+1)
  theta <- theta / as.vector(sqrt(crossprod(theta)))
  beta <- rnorm(n+1)
  beta <- beta / as.vector(sqrt(crossprod(beta)))
  beta.X <- rnorm(n+1)
  beta.X <- beta.X / as.vector(sqrt(crossprod(beta.X)))
  alpha <- rnorm(n+1, mean = mu.alpha, sd = sigma.alpha)

  X <- matrix(0, nrow = T+1, ncol = n+1)
  X[1, ] <- rnorm(n+1)
  Y <- matrix(0, nrow = T+1, ncol = n+1)
  Y[1, ] <- rnorm(n+1)

  for(i in 2:(T+1)){
    X[i, ] <- beta.X * X[i-1, ] + rnorm(n+1, sd = sigma.X)
    Y[i, ] <- eta + theta * Y[i-1, ] + beta * X[i, ] + 
      alpha * ifelse(i-1 > Tstar, 1, 0) + rnorm(n+1, sd = sigma) 
  }

  mu.alpha.hat <- rep(0, n)
  for(j in 2:(n+1)){
    m1 <- lm(Y[2:(T+1), j] ~ 1 + Y[1:T, j] + X[2:(T+1), j] + ifelse(1:T > Tstar,1,0))
    mu.alpha.hat[j-1] <- coefficients(m1)[4]
  }

  alpha.hat <- mean(mu.alpha.hat)
  #alpha.hat
  #sd(mu.alpha.hat) / sqrt(length(mu.alpha.hat))

  m1 <- lm(Y[2:(Tstar+1), 1] ~ 1 + Y[1:Tstar, 1] + X[2:(Tstar+1), 1])
  beta.hat <- coefficients(m1)
  Ypred.original <- as.vector(beta.hat %*% c(1, Y[Tstar+1,1], X[Tstar+2,1]))
  Ypred.adjusted <- Ypred.original + alpha.hat

  R.original <- (Ypred.original - Y[Tstar+2,1])^2
  R.adjusted <- (Ypred.adjusted - Y[Tstar+2,1])^2

  c(R.original, R.adjusted)
}
```

We set the random seed and declare the number of cores that our multicore implementation will use.

```{r}
ncores <- detectCores() - 1
registerDoParallel(cores = ncores)
set.seed(13)
RNGkind("L'Ecuyer-CMRG")
nsim <- 2e3
```

## Simple model simulations with normal errors

Combinations

```{r simsetup}
ns <- c(5, 1:4 * 10)
Ts <- c(1:4 * 10)
mus <- 1:5 / 2
sds <- 1:4
sim_params <- expand.grid(list(ns = ns, Ts = Ts, mus = mus, sds = sds))
```

```{r basicsims, cache = TRUE}
system.time({
  output_basic <- lapply(1:nrow(sim_params), FUN = function(j){
      n <- sim_params[j, 1]
      T <- sim_params[j, 2]
      mu.alpha <- sim_params[j, 3]
      sigma.alpha <- sim_params[j, 4]
      out <- foreach(i = 1:nsim, .combine = rbind) %dopar% {
        return(simstudy_normal(n = n, T = T, mu.alpha = mu.alpha, 
        sigma.alpha = sigma.alpha, sigma.X = 1, sigma = 1))
      }
      out
  })
})
```

```{r basicsims_ferror}
dat_basic <- cbind(sim_params, do.call(rbind, lapply(output_basic, colMeans)), 
  do.call(rbind, lapply(output_basic, function(x) colSds(x) / sqrt(nsim) )))
colnames(dat_basic)[5:8] <- c("mean_noadj", "mean_adj", "sd_noadj", "sd_adj") 
ggplot(dat_basic, aes(x = mus, y = mean_noadj - mean_adj, 
  group = as.factor(sds), color = as.factor(sds))) + 
  labs(title="Performance of adjustment via disparate information", 
    subtitle = "(rows increase in T, columns increase in n)",
    x ="shock means", y = "error without adjusting minus error after adjusting", 
    color = "shock sd") +
  geom_line() + geom_hline(yintercept = 0, color = "red") +
  theme_minimal() + 
  scale_x_continuous(breaks=c(1,2), minor_breaks = NULL) + 
  scale_y_continuous(minor_breaks = NULL) + 
  facet_grid(Ts ~ ns)
```

```{r, eval = FALSE, echo = FALSE}
n = 20 
T = 20
sigma.X = 1
sigma = 1
simstudy_gamma <- function(n = 20, T = 20, mu.alpha = 2, shape = 2, scale = 5, 
  neg = TRUE, sigma.X = 1, sigma = 1){

  Tstar <- floor(T/2)
  eta <- rnorm(n+1)
  theta <- rnorm(n+1)
  theta <- theta / as.vector(sqrt(crossprod(theta)))
  beta <- rnorm(n+1)
  beta <- beta / as.vector(sqrt(crossprod(beta)))
  beta.X <- rnorm(n+1)
  beta.X <- beta.X / as.vector(sqrt(crossprod(beta.X)))

  neg <- ifelse(neg, 1, -1)
  alpha <- mu.alpha + shape*scale*neg - 
    neg*rgamma(n = n+1, shape = shape, scale = scale)

  X <- matrix(0, nrow = T+1, ncol = n+1)
  X[1, ] <- rnorm(n+1)
  Y <- matrix(0, nrow = T+1, ncol = n+1)
  Y[1, ] <- rnorm(n+1)

  for(i in 2:(T+1)){
    X[i, ] <- beta.X * X[i-1, ] + rnorm(n+1, sd = sigma.X)
    Y[i, ] <- eta + theta * Y[i-1, ] + beta * X[i, ] + 
      alpha * ifelse(i-1 > Tstar, 1, 0) + rnorm(n+1, sd = sigma) 
  }

  mu.alpha.hat <- rep(0, n)
  for(j in 2:(n+1)){
    m1 <- lm(Y[2:(T+1), j] ~ 1 + Y[1:T, j] + X[2:(T+1), j] + ifelse(1:T > Tstar,1,0))
    mu.alpha.hat[j-1] <- coefficients(m1)[4]
  }

  alpha.hat <- mean(mu.alpha.hat)
  #alpha.hat
  #sd(mu.alpha.hat) / sqrt(length(mu.alpha.hat))

  m1 <- lm(Y[2:(Tstar+1), 1] ~ 1 + Y[1:Tstar, 1] + X[2:(Tstar+1), 1])
  beta.hat <- coefficients(m1)
  Ypred.original <- as.vector(beta.hat %*% c(1, Y[Tstar+1,1], X[Tstar+2,1]))
  Ypred.adjusted <- Ypred.original + alpha.hat

  R.original <- (Ypred.original - Y[Tstar+2,1])^2
  R.adjusted <- (Ypred.adjusted - Y[Tstar+2,1])^2

  c(R.original, R.adjusted)
}

test2 <- sapply(1:n.sim, FUN = function(j) 
  simstudy_gamma(n = 100, T = 100,  mu.alpha = 2, shape = 2, scale = 4, 
    neg = FALSE, sigma.X = 1, sigma = 1))
rowMeans(test2) # average forecast error for each forecast
rowSds(test2) / sqrt(n.sim) # Monte Carlo standard errors corresponding to the above averages

mean(apply(test2, 2, which.min) - 1) # proportion that adjusting wins
diffs2 <- apply(test2, 2, diff)

# average reduction achieved from adjustment
mean(diffs2)

# Monte Carlo standard error corresponding to the above average
sd(diffs2) / sqrt(n.sim) 
```