### 2020-05-20 manuscript/comp.tex
edit of literture review, parametric bootstrap, and Figure 1

% Created by tikzDevice version 0.12.3 on 2020-05-18 21:52:26
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,289.08);
\begin{scope}
\path[clip] ( 48.00, 48.00) rectangle (554.16,277.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 66.75,172.95) --
	( 68.63,152.47) --
	( 70.51,145.69) --
	( 72.39,140.03) --
	( 74.28,104.28) --
	( 76.16,151.54) --
	( 78.04,154.95) --
	( 79.92,133.81) --
	( 81.80,175.35) --
	( 83.69,166.88) --
	( 85.57,125.70) --
	( 87.45,170.76) --
	( 89.33,129.50) --
	( 91.22,171.48) --
	( 93.10,157.93) --
	( 94.98,132.20) --
	( 96.86,134.97) --
	( 98.74,121.20) --
	(100.63,132.31) --
	(102.51,167.74) --
	(104.39,139.43) --
	(106.27,161.95) --
	(108.15,118.76) --
	(110.04,173.04) --
	(111.92,101.84) --
	(113.80,132.79) --
	(115.68,156.33) --
	(117.57,121.16) --
	(119.45,173.74) --
	(121.33,100.01) --
	(123.21,122.32) --
	(125.09,136.35) --
	(126.98,148.45) --
	(128.86, 99.38) --
	(130.74,159.11) --
	(132.62,177.96) --
	(134.51,146.68) --
	(136.39, 99.65) --
	(138.27,141.14) --
	(140.15,113.97) --
	(142.03, 56.48) --
	(143.92,137.78) --
	(145.80,157.07) --
	(147.68,135.25) --
	(149.56,140.67) --
	(151.45,155.40) --
	(153.33,144.28) --
	(155.21,109.64) --
	(157.09,161.65) --
	(158.97,120.57) --
	(160.86,134.53) --
	(162.74,187.79) --
	(164.62,127.58) --
	(166.50,140.39) --
	(168.39,128.19) --
	(170.27,170.09) --
	(172.15,121.23) --
	(174.03,147.74) --
	(175.91,166.20) --
	(177.80,152.19) --
	(179.68, 95.16) --
	(181.56,137.35) --
	(183.44,180.08) --
	(185.32,152.17) --
	(187.21,170.34) --
	(189.09,168.84) --
	(190.97,130.81) --
	(192.85, 95.27) --
	(194.74,162.93) --
	(196.62,138.78) --
	(198.50,138.61) --
	(200.38,173.02) --
	(202.26,167.54) --
	(204.15,162.63) --
	(206.03,148.90) --
	(207.91,177.00) --
	(209.79,135.59) --
	(211.68,142.01) --
	(213.56,146.72) --
	(215.44,147.34) --
	(217.32,137.83) --
	(219.20,161.38) --
	(221.09,135.23) --
	(222.97,113.76) --
	(224.85,132.15) --
	(226.73,152.10) --
	(228.62,133.70) --
	(230.50,142.92) --
	(232.38,137.94) --
	(234.26, 80.23) --
	(236.14,128.05) --
	(238.03,140.12) --
	(239.91,113.33) --
	(241.79,135.32) --
	(243.67,145.57) --
	(245.56,148.21) --
	(247.44,181.13) --
	(249.32,123.87) --
	(251.20,167.03) --
	(253.08,176.92) --
	(254.97,146.67) --
	(256.85,137.49) --
	(258.73,148.71) --
	(260.61,158.20) --
	(262.49,144.01) --
	(264.38,140.56) --
	(266.26,134.50) --
	(268.14,164.72) --
	(270.02,113.97) --
	(271.91,173.50) --
	(273.79,131.97) --
	(275.67,135.32) --
	(277.55,143.59) --
	(279.43,134.00) --
	(281.32,182.28) --
	(283.20,134.50) --
	(285.08,136.30) --
	(286.96,121.98) --
	(288.85,170.76) --
	(290.73,168.83) --
	(292.61,130.32) --
	(294.49,130.95) --
	(296.37,167.83) --
	(298.26,106.18) --
	(300.14,158.89) --
	(302.02,221.32) --
	(303.90,262.43) --
	(305.79,181.52) --
	(307.67,221.00) --
	(309.55,219.36) --
	(311.43,223.97) --
	(313.31,225.88) --
	(315.20,172.43) --
	(317.08,211.80) --
	(318.96,229.45) --
	(320.84,222.24) --
	(322.73,252.46) --
	(324.61,190.77) --
	(326.49,224.11) --
	(328.37,194.35) --
	(330.25,184.33) --
	(332.14,175.51) --
	(334.02,208.94) --
	(335.90,208.16) --
	(337.78,210.48) --
	(339.67,204.77) --
	(341.55,200.42) --
	(343.43,207.20) --
	(345.31,194.28) --
	(347.19,175.97) --
	(349.08,221.82) --
	(350.96,230.52) --
	(352.84,206.54) --
	(354.72,213.96) --
	(356.60,241.92) --
	(358.49,206.43) --
	(360.37,254.17) --
	(362.25,205.63) --
	(364.13,199.30) --
	(366.02,207.87) --
	(367.90,210.32) --
	(369.78,198.03) --
	(371.66,248.03) --
	(373.54,206.39) --
	(375.43,246.52) --
	(377.31,210.82) --
	(379.19,232.38) --
	(381.07,213.29) --
	(382.96,237.01) --
	(384.84,208.79) --
	(386.72,192.52) --
	(388.60,236.32) --
	(390.48,175.56) --
	(392.37,250.12) --
	(394.25,185.69) --
	(396.13,231.06) --
	(398.01,191.78) --
	(399.90,226.76) --
	(401.78,199.54) --
	(403.66,165.06) --
	(405.54,255.85) --
	(407.42,210.34) --
	(409.31,199.91) --
	(411.19,229.86) --
	(413.07,203.76) --
	(414.95,231.64) --
	(416.84,230.77) --
	(418.72,190.26) --
	(420.60,203.84) --
	(422.48,209.53) --
	(424.36,220.43) --
	(426.25,173.24) --
	(428.13,209.84) --
	(430.01,159.64) --
	(431.89,251.30) --
	(433.77,197.02) --
	(435.66,216.20) --
	(437.54,224.42) --
	(439.42,193.29) --
	(441.30,213.98) --
	(443.19,172.72) --
	(445.07,245.00) --
	(446.95,192.05) --
	(448.83,235.97) --
	(450.71,204.00) --
	(452.60,164.26) --
	(454.48,238.03) --
	(456.36,185.07) --
	(458.24,227.07) --
	(460.13,222.82) --
	(462.01,212.73) --
	(463.89,268.60) --
	(465.77,192.34) --
	(467.65,239.77) --
	(469.54,202.33) --
	(471.42,231.75) --
	(473.30,220.58) --
	(475.18,187.28) --
	(477.07,206.99) --
	(478.95,199.08) --
	(480.83,226.56) --
	(482.71,221.75) --
	(484.59,227.59) --
	(486.48,203.00) --
	(488.36,227.21) --
	(490.24,209.56) --
	(492.12,221.87) --
	(494.01,245.38) --
	(495.89,236.19) --
	(497.77,241.22) --
	(499.65,231.39) --
	(501.53,253.76) --
	(503.42,191.19) --
	(505.30,203.77) --
	(507.18,247.46) --
	(509.06,219.46) --
	(510.94,214.21) --
	(512.83,191.34) --
	(514.71,248.79) --
	(516.59,197.34) --
	(518.47,220.12) --
	(520.36,259.69) --
	(522.24,240.39) --
	(524.12,221.17) --
	(526.00,214.51) --
	(527.88,259.99) --
	(529.77,220.91) --
	(531.65,215.72) --
	(533.53,233.07) --
	(535.41,219.94);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.98, 48.00) -- (533.53, 48.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.98, 48.00) -- ( 62.98, 42.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (157.09, 48.00) -- (157.09, 42.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (251.20, 48.00) -- (251.20, 42.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.31, 48.00) -- (345.31, 42.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (439.42, 48.00) -- (439.42, 42.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (533.53, 48.00) -- (533.53, 42.00);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 62.98, 26.40) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (157.09, 26.40) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (251.20, 26.40) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (345.31, 26.40) {150};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (439.42, 26.40) {200};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (533.53, 26.40) {250};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 55.58) -- ( 48.00,270.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 55.58) -- ( 42.00, 55.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 98.61) -- ( 42.00, 98.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,141.64) -- ( 42.00,141.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,184.67) -- ( 42.00,184.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,227.69) -- ( 42.00,227.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,270.72) -- ( 42.00,270.72);

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 33.60, 55.58) {$-400$};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 33.60, 98.61) {$-200$};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 33.60,141.64) {0};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 33.60,184.67) {200};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 33.60,227.69) {400};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 33.60,270.72) {600};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 48.00) --
	(554.16, 48.00) --
	(554.16,277.08) --
	( 48.00,277.08) --
	( 48.00, 48.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,289.08);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (301.08,  2.40) {Time $T$};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (  9.60,162.54) {Response};
\end{scope}
\begin{scope}
\path[clip] ( 48.00, 48.00) rectangle (554.16,277.08);
\definecolor{drawColor}{RGB}{255,106,106}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] ( 66.75,142.83) --
	( 68.63,140.13) --
	( 70.51,141.93) --
	( 72.39,141.91) --
	( 74.28,142.43) --
	( 76.16,146.39) --
	( 78.04,141.61) --
	( 79.92,141.14) --
	( 81.80,143.78) --
	( 83.69,140.06) --
	( 85.57,141.28) --
	( 87.45,144.25) --
	( 89.33,140.02) --
	( 91.22,143.17) --
	( 93.10,140.45) --
	( 94.98,141.62) --
	( 96.86,143.92) --
	( 98.74,143.18) --
	(100.63,144.99) --
	(102.51,143.92) --
	(104.39,140.49) --
	(106.27,143.03) --
	(108.15,141.75) --
	(110.04,144.93) --
	(111.92,140.19) --
	(113.80,146.77) --
	(115.68,143.66) --
	(117.57,141.45) --
	(119.45,144.73) --
	(121.33,140.27) --
	(123.21,146.15) --
	(125.09,144.70) --
	(126.98,143.27) --
	(128.86,141.92) --
	(130.74,146.15) --
	(132.62,140.59) --
	(134.51,139.83) --
	(136.39,141.78) --
	(138.27,147.01) --
	(140.15,142.91) --
	(142.03,145.16) --
	(143.92,150.29) --
	(145.80,143.35) --
	(147.68,141.31) --
	(149.56,142.82) --
	(151.45,142.93) --
	(153.33,141.23) --
	(155.21,142.28) --
	(157.09,144.73) --
	(158.97,141.28) --
	(160.86,144.80) --
	(162.74,143.76) --
	(164.62,139.14) --
	(166.50,144.55) --
	(168.39,142.11) --
	(170.27,143.40) --
	(172.15,140.31) --
	(174.03,144.97) --
	(175.91,141.97) --
	(177.80,140.53) --
	(179.68,141.81) --
	(181.56,146.17) --
	(183.44,143.43) --
	(185.32,139.06) --
	(187.21,142.38) --
	(189.09,140.37) --
	(190.97,141.03) --
	(192.85,143.88) --
	(194.74,146.77) --
	(196.62,140.46) --
	(198.50,143.03) --
	(200.38,141.89) --
	(202.26,140.44) --
	(204.15,139.71) --
	(206.03,141.07) --
	(207.91,142.19) --
	(209.79,139.86) --
	(211.68,142.89) --
	(213.56,142.17) --
	(215.44,142.86) --
	(217.32,142.40) --
	(219.20,143.71) --
	(221.09,141.02) --
	(222.97,142.74) --
	(224.85,145.76) --
	(226.73,143.65) --
	(228.62,141.79) --
	(230.50,143.50) --
	(232.38,142.42) --
	(234.26,142.91) --
	(236.14,147.76) --
	(238.03,143.33) --
	(239.91,142.66) --
	(241.79,145.02) --
	(243.67,143.33) --
	(245.56,142.23) --
	(247.44,142.51) --
	(249.32,139.83) --
	(251.20,143.91) --
	(253.08,140.91) --
	(254.97,139.48) --
	(256.85,142.03) --
	(258.73,142.30) --
	(260.61,142.42) --
	(262.49,141.41) --
	(264.38,142.28) --
	(266.26,143.37) --
	(268.14,143.80) --
	(270.02,141.46) --
	(271.91,145.00) --
	(273.79,140.14) --
	(275.67,144.39) --
	(277.55,142.54) --
	(279.43,143.15) --
	(281.32,143.33) --
	(283.20,139.76) --
	(285.08,143.03) --
	(286.96,142.39) --
	(288.85,144.66) --
	(290.73,140.34) --
	(292.61,139.93) --
	(294.49,143.71) --
	(296.37,143.28) --
	(298.26,140.41) --
	(300.14,145.43) --
	(302.02,140.85) --
	(303.90,136.75) --
	(305.79,132.94) --
	(307.67,139.42) --
	(309.55,135.87) --
	(311.43,136.63) --
	(313.31,135.78) --
	(315.20,135.92) --
	(317.08,140.12) --
	(318.96,136.56) --
	(320.84,134.76) --
	(322.73,135.54) --
	(324.61,133.98) --
	(326.49,138.52) --
	(328.37,135.72) --
	(330.25,138.98) --
	(332.14,139.33) --
	(334.02,140.22) --
	(335.90,136.88) --
	(337.78,137.21) --
	(339.67,136.66) --
	(341.55,137.15) --
	(343.43,138.26) --
	(345.31,136.84) --
	(347.19,138.24) --
	(349.08,139.43) --
	(350.96,136.01) --
	(352.84,135.41) --
	(354.72,136.89) --
	(356.60,136.87) --
	(358.49,134.00) --
	(360.37,138.00) --
	(362.25,133.77) --
	(364.13,137.44) --
	(366.02,136.97) --
	(367.90,136.64) --
	(369.78,136.51) --
	(371.66,138.07) --
	(373.54,134.02) --
	(375.43,137.34) --
	(377.31,134.24) --
	(379.19,137.50) --
	(381.07,134.80) --
	(382.96,138.20) --
	(384.84,133.93) --
	(386.72,136.89) --
	(388.60,138.15) --
	(390.48,134.68) --
	(392.37,139.28) --
	(394.25,133.59) --
	(396.13,139.33) --
	(398.01,135.30) --
	(399.90,137.96) --
	(401.78,135.47) --
	(403.66,137.90) --
	(405.54,140.60) --
	(407.42,133.45) --
	(409.31,136.69) --
	(411.19,137.22) --
	(413.07,135.30) --
	(414.95,137.11) --
	(416.84,134.45) --
	(418.72,135.26) --
	(420.60,139.07) --
	(422.48,137.43) --
	(424.36,137.15) --
	(426.25,136.50) --
	(428.13,140.82) --
	(430.01,138.00) --
	(431.89,141.05) --
	(433.77,133.05) --
	(435.66,137.92) --
	(437.54,136.77) --
	(439.42,135.43) --
	(441.30,139.49) --
	(443.19,136.51) --
	(445.07,140.20) --
	(446.95,134.28) --
	(448.83,139.11) --
	(450.71,134.98) --
	(452.60,137.08) --
	(454.48,141.67) --
	(456.36,134.46) --
	(458.24,139.63) --
	(460.13,134.69) --
	(462.01,135.78) --
	(463.89,137.05) --
	(465.77,131.53) --
	(467.65,138.29) --
	(469.54,134.65) --
	(471.42,137.65) --
	(473.30,135.03) --
	(475.18,136.76) --
	(477.07,138.74) --
	(478.95,137.36) --
	(480.83,138.23) --
	(482.71,135.82) --
	(484.59,135.95) --
	(486.48,135.89) --
	(488.36,136.74) --
	(490.24,135.40) --
	(492.12,136.77) --
	(494.01,135.94) --
	(495.89,133.46) --
	(497.77,134.48) --
	(499.65,134.68) --
	(501.53,135.56) --
	(503.42,134.09) --
	(505.30,139.32) --
	(507.18,138.00) --
	(509.06,133.98) --
	(510.94,136.16) --
	(512.83,136.35) --
	(514.71,138.27) --
	(516.59,133.84) --
	(518.47,137.50) --
	(520.36,136.90) --
	(522.24,132.66) --
	(524.12,134.57) --
	(526.00,136.61) --
	(527.88,136.86) --
	(529.77,132.70) --
	(531.65,135.88) --
	(533.53,136.61) --
	(535.41,134.54);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt ,line join=round,line cap=round] ( 66.75,142.83) --
	( 68.63,140.13) --
	( 70.51,141.93) --
	( 72.39,141.91) --
	( 74.28,142.43) --
	( 76.16,146.39) --
	( 78.04,141.61) --
	( 79.92,141.14) --
	( 81.80,143.78) --
	( 83.69,140.06) --
	( 85.57,141.28) --
	( 87.45,144.25) --
	( 89.33,140.02) --
	( 91.22,143.17) --
	( 93.10,140.45) --
	( 94.98,141.62) --
	( 96.86,143.92) --
	( 98.74,143.18) --
	(100.63,144.99) --
	(102.51,143.92) --
	(104.39,140.49) --
	(106.27,143.03) --
	(108.15,141.75) --
	(110.04,144.93) --
	(111.92,140.19) --
	(113.80,146.77) --
	(115.68,143.66) --
	(117.57,141.45) --
	(119.45,144.73) --
	(121.33,140.27) --
	(123.21,146.15) --
	(125.09,144.70) --
	(126.98,143.27) --
	(128.86,141.92) --
	(130.74,146.15) --
	(132.62,140.59) --
	(134.51,139.83) --
	(136.39,141.78) --
	(138.27,147.01) --
	(140.15,142.91) --
	(142.03,145.16) --
	(143.92,150.29) --
	(145.80,143.35) --
	(147.68,141.31) --
	(149.56,142.82) --
	(151.45,142.93) --
	(153.33,141.23) --
	(155.21,142.28) --
	(157.09,144.73) --
	(158.97,141.28) --
	(160.86,144.80) --
	(162.74,143.76) --
	(164.62,139.14) --
	(166.50,144.55) --
	(168.39,142.11) --
	(170.27,143.40) --
	(172.15,140.31) --
	(174.03,144.97) --
	(175.91,141.97) --
	(177.80,140.53) --
	(179.68,141.81) --
	(181.56,146.17) --
	(183.44,143.43) --
	(185.32,139.06) --
	(187.21,142.38) --
	(189.09,140.37) --
	(190.97,141.03) --
	(192.85,143.88) --
	(194.74,146.77) --
	(196.62,140.46) --
	(198.50,143.03) --
	(200.38,141.89) --
	(202.26,140.44) --
	(204.15,139.71) --
	(206.03,141.07) --
	(207.91,142.19) --
	(209.79,139.86) --
	(211.68,142.89) --
	(213.56,142.17) --
	(215.44,142.86) --
	(217.32,142.40) --
	(219.20,143.71) --
	(221.09,141.02) --
	(222.97,142.74) --
	(224.85,145.76) --
	(226.73,143.65) --
	(228.62,141.79) --
	(230.50,143.50) --
	(232.38,142.42) --
	(234.26,142.91) --
	(236.14,147.76) --
	(238.03,143.33) --
	(239.91,142.66) --
	(241.79,145.02) --
	(243.67,143.33) --
	(245.56,142.23) --
	(247.44,142.51) --
	(249.32,139.83) --
	(251.20,143.91) --
	(253.08,140.91) --
	(254.97,139.48) --
	(256.85,142.03) --
	(258.73,142.30) --
	(260.61,142.42) --
	(262.49,141.41) --
	(264.38,142.28) --
	(266.26,143.37) --
	(268.14,143.80) --
	(270.02,141.46) --
	(271.91,145.00) --
	(273.79,140.14) --
	(275.67,144.39) --
	(277.55,142.54) --
	(279.43,143.15) --
	(281.32,143.33) --
	(283.20,139.76) --
	(285.08,143.03) --
	(286.96,142.39) --
	(288.85,144.66) --
	(290.73,140.34) --
	(292.61,139.93) --
	(294.49,143.71) --
	(296.37,143.28) --
	(298.26,140.41) --
	(300.14,194.20) --
	(302.02,189.62) --
	(303.90,185.53) --
	(305.79,181.72) --
	(307.67,188.20) --
	(309.55,184.65) --
	(311.43,185.40) --
	(313.31,184.56) --
	(315.20,184.70) --
	(317.08,188.90) --
	(318.96,185.34) --
	(320.84,183.53) --
	(322.73,184.32) --
	(324.61,182.76) --
	(326.49,187.29) --
	(328.37,184.50) --
	(330.25,187.76) --
	(332.14,188.11) --
	(334.02,189.00) --
	(335.90,185.66) --
	(337.78,185.99) --
	(339.67,185.43) --
	(341.55,185.92) --
	(343.43,187.03) --
	(345.31,185.62) --
	(347.19,187.02) --
	(349.08,188.21) --
	(350.96,184.78) --
	(352.84,184.18) --
	(354.72,185.67) --
	(356.60,185.65) --
	(358.49,182.77) --
	(360.37,186.77) --
	(362.25,182.54) --
	(364.13,186.22) --
	(366.02,185.75) --
	(367.90,185.41) --
	(369.78,185.28) --
	(371.66,186.84) --
	(373.54,182.80) --
	(375.43,186.12) --
	(377.31,183.02) --
	(379.19,186.28) --
	(381.07,183.58) --
	(382.96,186.97) --
	(384.84,182.70) --
	(386.72,185.67) --
	(388.60,186.93) --
	(390.48,183.46) --
	(392.37,188.06) --
	(394.25,182.37) --
	(396.13,188.11) --
	(398.01,184.08) --
	(399.90,186.74) --
	(401.78,184.24) --
	(403.66,186.67) --
	(405.54,189.37) --
	(407.42,182.23) --
	(409.31,185.46) --
	(411.19,186.00) --
	(413.07,184.08) --
	(414.95,185.89) --
	(416.84,183.23) --
	(418.72,184.04) --
	(420.60,187.85) --
	(422.48,186.21) --
	(424.36,185.93) --
	(426.25,185.27) --
	(428.13,189.59) --
	(430.01,186.78) --
	(431.89,189.83) --
	(433.77,181.83) --
	(435.66,186.69) --
	(437.54,185.54) --
	(439.42,184.20) --
	(441.30,188.26) --
	(443.19,185.29) --
	(445.07,188.98) --
	(446.95,183.06) --
	(448.83,187.89) --
	(450.71,183.76) --
	(452.60,185.86) --
	(454.48,190.44) --
	(456.36,183.23) --
	(458.24,188.40) --
	(460.13,183.47) --
	(462.01,184.56) --
	(463.89,185.83) --
	(465.77,180.31) --
	(467.65,187.07) --
	(469.54,183.42) --
	(471.42,186.43) --
	(473.30,183.81) --
	(475.18,185.53) --
	(477.07,187.52) --
	(478.95,186.14) --
	(480.83,187.00) --
	(482.71,184.60) --
	(484.59,184.73) --
	(486.48,184.66) --
	(488.36,185.52) --
	(490.24,184.18) --
	(492.12,185.54) --
	(494.01,184.71) --
	(495.89,182.24) --
	(497.77,183.25) --
	(499.65,183.46) --
	(501.53,184.33) --
	(503.42,182.87) --
	(505.30,188.10) --
	(507.18,186.77) --
	(509.06,182.75) --
	(510.94,184.94) --
	(512.83,185.13) --
	(514.71,187.05) --
	(516.59,182.61) --
	(518.47,186.28) --
	(520.36,185.68) --
	(522.24,181.44) --
	(524.12,183.35) --
	(526.00,185.39) --
	(527.88,185.63) --
	(529.77,181.48) --
	(531.65,184.66) --
	(533.53,185.38) --
	(535.41,183.31);
\definecolor{drawColor}{RGB}{0,191,255}

\path[draw=drawColor,line width= 0.8pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (298.26, 55.58) -- (298.26,270.72);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,196.47) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,189.77) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,206.04) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,189.73) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,168.04) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,154.00) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,222.37) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,169.47) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,198.29) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,201.79) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,175.89) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,186.08) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,164.51) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,212.93) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,202.61) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,181.94) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,162.52) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,221.82) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,193.57) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,216.81) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,190.10) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,176.09) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,192.35) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,184.10) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,224.01) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,181.45) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,185.50) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,188.50) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,187.45) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,186.29) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,184.69) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,227.02) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,193.52) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,191.68) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,169.72) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,192.50) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,195.71) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,184.13) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,192.57) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.26,174.59) circle (  2.25);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (320.84,137.34) -- (320.84,186.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (318.31,141.72) --
	(320.84,137.34) --
	(323.37,141.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (323.37,181.74) --
	(320.84,186.12) --
	(318.31,181.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (277.55,190.42) -- (292.61,190.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (288.23,187.89) --
	(292.61,190.42) --
	(288.23,192.94);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (373.54,158.80) {$\displaystyle \hat{\alpha} = \frac{1}{n}\sum_{i=1}^{n} \hat{\alpha}_i=226.72$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (270.02,187.92) {$\hat{\alpha}_i$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (354.72, 63.76) {shock time point $T^*_1$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.98,270.72) rectangle (206.60,234.72);
\definecolor{drawColor}{RGB}{255,106,106}

\path[draw=drawColor,line width= 0.8pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] ( 71.98,258.72) -- ( 89.98,258.72);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.8pt,dash pattern=on 1pt off 3pt ,line join=round,line cap=round] ( 71.98,246.72) -- ( 89.98,246.72);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 98.98,255.28) {$\hat{Y}$ without shock effects};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 98.98,243.28) {$\hat{Y}$ with shock effects};
\end{scope}
\end{tikzpicture}

\end{document}