### 2020-06-06 manuscript/fig2.tex
updated data analysis, discussions, tables for parametric bootstrap

% Created by tikzDevice version 0.12.3 on 2020-06-05 08:57:46
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (505.89,361.35);
\begin{scope}
\path[clip] ( 31.71, 30.69) rectangle (500.39,338.69);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.71, 53.81) --
	(500.39, 53.81);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.71,151.88) --
	(500.39,151.88);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.71,249.96) --
	(500.39,249.96);

\path[draw=drawColor,line width= 0.3pt,line join=round] (121.74, 30.69) --
	(121.74,338.69);

\path[draw=drawColor,line width= 0.3pt,line join=round] (259.18, 30.69) --
	(259.18,338.69);

\path[draw=drawColor,line width= 0.3pt,line join=round] (396.62, 30.69) --
	(396.62,338.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.71,102.85) --
	(500.39,102.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.71,200.92) --
	(500.39,200.92);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.71,299.00) --
	(500.39,299.00);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 53.02, 30.69) --
	( 53.02,338.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (190.46, 30.69) --
	(190.46,338.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (327.90, 30.69) --
	(327.90,338.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] (465.34, 30.69) --
	(465.34,338.69);
\definecolor{drawColor}{RGB}{0,0,0}
\definecolor{fillColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 66.76,324.69) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 80.50,309.20) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] ( 94.25,310.57) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (107.99,303.80) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (121.74,307.73) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (135.48,293.41) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (149.23,290.95) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (162.97,264.57) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (176.71,287.03) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (190.46,283.40) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (204.20,279.28) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (217.95,278.11) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (231.69,286.34) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (245.43,296.64) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (259.18,287.42) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (272.92,285.46) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (286.67,282.32) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (300.41,291.74) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (314.16,288.01) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (327.90,283.70) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (341.64,263.49) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (355.39,238.48) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (369.13,212.40) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (382.88,172.77) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (396.62,185.42) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (410.37,193.37) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (424.11,182.48) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (437.85,195.33) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (451.60,178.36) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (465.34,155.12) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (479.09, 44.69) circle (  1.96);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (479.09, 91.63) --
	(481.73, 87.05) --
	(476.44, 87.05) --
	(479.09, 91.63);
\definecolor{drawColor}{RGB}{0,191,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (476.31, 91.18) -- (481.86, 91.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (479.09, 88.41) -- (479.09, 93.96);
\definecolor{drawColor}{RGB}{238,99,99}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.12, 86.20) -- (481.05, 90.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.12, 90.12) -- (481.05, 86.20);
\definecolor{drawColor}{RGB}{238,130,238}
\definecolor{fillColor}{RGB}{238,130,238}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round,fill=fillColor] (479.09,141.49) circle (  1.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 66.76,313.70) --
	( 80.50,313.67) --
	( 94.25,302.10) --
	(107.99,304.27) --
	(121.74,297.06) --
	(135.48,299.42) --
	(149.23,290.97) --
	(162.97,283.99) --
	(176.71,258.33) --
	(190.46,290.13) --
	(204.20,281.01) --
	(217.95,276.50) --
	(231.69,274.52) --
	(245.43,289.86) --
	(259.18,300.87) --
	(272.92,286.25) --
	(286.67,282.81) --
	(300.41,276.99) --
	(314.16,288.22) --
	(327.90,280.06) --
	(341.64,273.21) --
	(355.39,252.42) --
	(369.13,229.35) --
	(382.88,199.60) --
	(396.62,163.01) --
	(410.37,175.15) --
	(424.11,190.38) --
	(437.85,181.33) --
	(451.60,180.97) --
	(465.34,175.47) --
	(479.09,141.49);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 66.76,148.08) {$\hat{y}_{1, T_1^* + 1}^{1}$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 66.76,131.74) {$\hat{y}_{1, T_1^* + 1}^{1} + \hat{\alpha}_{\rm adj}$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 66.76,115.39) {$\hat{y}_{1, T_1^* + 1}^{1} + \hat{\alpha}_{\rm IVW}$};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 66.76, 99.04) {$\hat{y}_{1, T_1^* + 1}^{1} + \hat{\alpha}_{\rm wadj}$};

\path[fill=fillColor] ( 53.02,151.88) circle (  1.96);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 53.02,138.59) --
	( 55.66,134.01) --
	( 50.37,134.01) --
	( 53.02,138.59);
\definecolor{drawColor}{RGB}{0,191,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 50.24,119.19) -- ( 55.79,119.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 53.02,116.42) -- ( 53.02,121.97);
\definecolor{drawColor}{RGB}{238,99,99}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 51.05,100.88) -- ( 54.98,104.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 51.05,104.81) -- ( 54.98,100.88);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,361.35);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 26.76, 99.81) {40};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 26.76,197.89) {50};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 26.76,295.97) {60};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,361.35);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at ( 53.02, 19.68) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (190.46, 19.68) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (327.90, 19.68) {20};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.88] at (465.34, 19.68) {30};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at (266.05,  7.64) {Day};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.10] at ( 13.08,184.69) {Closing Stock price (in USD)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,361.35);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.32] at (266.05,346.76) {Conoco Phillips Stock Forecasting (2020--01--24 to 2020--03--09)};
\end{scope}
\end{tikzpicture}

\end{document}