### 2020-08-05 manuscript/comp.tex

% Created by tikzDevice version 0.12.3.1 on 2020-08-05 11:58:05
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,252.94);
\begin{scope}
\path[clip] ( 48.00, 48.00) rectangle (554.16,240.94);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 72.43,141.91) --
	( 75.27,148.65) --
	( 78.11,138.04) --
	( 80.95,120.32) --
	( 83.79,144.55) --
	( 86.63,128.04) --
	( 89.47,158.74) --
	( 92.31,142.28) --
	( 95.15,144.24) --
	( 97.99,150.16) --
	(100.83,129.86) --
	(103.67,156.20) --
	(106.51,161.19) --
	(109.35,150.66) --
	(112.19,143.27) --
	(115.03,126.88) --
	(117.87,118.05) --
	(120.71,139.15) --
	(123.55,135.40) --
	(126.40,140.74) --
	(129.24,149.36) --
	(132.08,115.48) --
	(134.92,139.71) --
	(137.76,114.03) --
	(140.60,125.67) --
	(143.44,139.07) --
	(146.28,144.69) --
	(149.12,130.89) --
	(151.96,140.09) --
	(154.80,126.64) --
	(157.64,111.23) --
	(160.48,121.82) --
	(163.32,149.80) --
	(166.16,149.85) --
	(169.00,139.78) --
	(171.84,158.01) --
	(174.68,141.87) --
	(177.52,133.42) --
	(180.36,131.05) --
	(183.20,129.40) --
	(186.04,142.69) --
	(188.88,144.63) --
	(191.72,130.06) --
	(194.56,121.56) --
	(197.41,118.55) --
	(200.25,119.73) --
	(203.09,140.14) --
	(205.93,137.20) --
	(208.77,125.79) --
	(211.61,150.02) --
	(214.45,134.16) --
	(217.29,137.62) --
	(220.13,143.96) --
	(222.97,145.02) --
	(225.81,137.14) --
	(228.65,147.61) --
	(231.49,162.41) --
	(234.33,138.41) --
	(237.17,162.38) --
	(240.01,127.19) --
	(242.85,125.12) --
	(245.69,134.60) --
	(248.53,134.89) --
	(251.37,158.83) --
	(254.21,122.69) --
	(257.05,117.16) --
	(259.89,147.80) --
	(262.73,138.62) --
	(265.57,123.35) --
	(268.42,140.39) --
	(271.26,137.29) --
	(274.10,148.46) --
	(276.94,135.80) --
	(279.78,134.54) --
	(282.62,139.54) --
	(285.46,143.03) --
	(288.30,143.37) --
	(291.14,107.15) --
	(293.98,121.09) --
	(296.82,144.70) --
	(299.66,143.32) --
	(302.50,149.80) --
	(305.34,113.62) --
	(308.18,139.96) --
	(311.02,134.15) --
	(313.86,130.85) --
	(316.70,135.42) --
	(319.54,131.11) --
	(322.38,145.18) --
	(325.22,143.55) --
	(328.06,141.15) --
	(330.90,125.31) --
	(333.74,129.29) --
	(336.59,118.52) --
	(339.43,146.06) --
	(342.27,140.93) --
	(345.11,151.97) --
	(347.95,136.60) --
	(350.79,114.01) --
	(353.63,120.33) --
	(356.47,142.22) --
	(359.31,130.73) --
	(362.15,138.80) --
	(364.99,114.92) --
	(367.83,152.60) --
	(370.67,139.26) --
	(373.51,137.13) --
	(376.35,140.56) --
	(379.19,149.07) --
	(382.03,151.38) --
	(384.87,153.12) --
	(387.71,126.88) --
	(390.55,137.55) --
	(393.39,137.71) --
	(396.23,129.83) --
	(399.07,123.04) --
	(401.91,117.89) --
	(404.75,132.35) --
	(407.60,141.10) --
	(410.44,137.00) --
	(413.28,143.73) --
	(416.12,131.05) --
	(418.96,129.54) --
	(421.80,129.16) --
	(424.64,128.38) --
	(427.48,164.19);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.75, 48.00) -- (492.81, 48.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.75, 48.00) -- ( 66.75, 42.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (208.77, 48.00) -- (208.77, 42.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (350.79, 48.00) -- (350.79, 42.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (492.81, 48.00) -- (492.81, 42.00);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 66.75, 26.40) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (208.77, 26.40) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (350.79, 26.40) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (492.81, 26.40) {150};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 67.43) -- ( 48.00,190.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 67.43) -- ( 42.00, 67.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,128.99) -- ( 42.00,128.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,190.54) -- ( 42.00,190.54);

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 33.60, 67.43) {-5};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 33.60,128.99) {0};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 33.60,190.54) {5};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 48.00) --
	(554.16, 48.00) --
	(554.16,240.94) --
	( 48.00,240.94) --
	( 48.00, 48.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (301.08,  2.40) {Time $T$};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (  9.60,144.47) {Response};
\end{scope}
\begin{scope}
\path[clip] ( 48.00, 48.00) rectangle (554.16,240.94);
\definecolor{drawColor}{RGB}{255,106,106}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] ( 72.43,137.81) --
	( 75.27,136.65) --
	( 78.11,143.49) --
	( 80.95,137.43) --
	( 83.79,130.64) --
	( 86.63,140.10) --
	( 89.47,133.95) --
	( 92.31,138.95) --
	( 95.15,138.13) --
	( 97.99,134.38) --
	(100.83,138.70) --
	(103.67,136.65) --
	(106.51,142.33) --
	(109.35,140.08) --
	(112.19,140.76) --
	(115.03,135.30) --
	(117.87,131.53) --
	(120.71,134.31) --
	(123.55,136.31) --
	(126.40,138.07) --
	(129.24,138.68) --
	(132.08,133.35) --
	(134.92,135.06) --
	(137.76,137.42) --
	(140.60,132.08) --
	(143.44,134.84) --
	(146.28,141.15) --
	(149.12,136.50) --
	(151.96,137.75) --
	(154.80,136.26) --
	(157.64,132.36) --
	(160.48,135.15) --
	(163.32,133.44) --
	(166.16,135.02) --
	(169.00,137.46) --
	(171.84,137.44) --
	(174.68,138.85) --
	(177.52,137.64) --
	(180.36,136.64) --
	(183.20,136.98) --
	(186.04,135.19) --
	(188.88,135.00) --
	(191.72,137.46) --
	(194.56,133.62) --
	(197.41,136.58) --
	(200.25,133.35) --
	(203.09,137.85) --
	(205.93,138.76) --
	(208.77,132.56) --
	(211.61,136.25) --
	(214.45,135.48) --
	(217.29,136.00) --
	(220.13,134.15) --
	(222.97,138.92) --
	(225.81,140.34) --
	(228.65,137.72) --
	(231.49,138.74) --
	(234.33,140.75) --
	(237.17,137.08) --
	(240.01,139.13) --
	(242.85,132.23) --
	(245.69,135.17) --
	(248.53,135.31) --
	(251.37,134.56) --
	(254.21,137.20) --
	(257.05,135.96) --
	(259.89,133.73) --
	(262.73,140.13) --
	(265.57,133.33) --
	(268.42,134.06) --
	(271.26,136.71) --
	(274.10,138.64) --
	(276.94,143.32) --
	(279.78,136.35) --
	(282.62,136.03) --
	(285.46,138.94) --
	(288.30,139.04) --
	(291.14,138.48) --
	(293.98,133.60) --
	(296.82,137.06) --
	(299.66,135.32) --
	(302.50,136.59) --
	(305.34,139.19) --
	(308.18,132.33) --
	(311.02,130.50) --
	(313.86,133.96) --
	(316.70,137.48) --
	(319.54,136.37) --
	(322.38,135.54) --
	(325.22,141.95) --
	(328.06,137.28) --
	(330.90,136.24) --
	(333.74,135.82) --
	(336.59,129.34) --
	(339.43,131.63) --
	(342.27,137.63) --
	(345.11,139.18) --
	(347.95,141.59) --
	(350.79,135.78) --
	(353.63,132.71) --
	(356.47,132.87) --
	(359.31,137.75) --
	(362.15,135.79) --
	(364.99,140.06) --
	(367.83,133.34) --
	(370.67,138.30) --
	(373.51,137.87) --
	(376.35,138.13) --
	(379.19,134.63) --
	(382.03,136.24) --
	(384.87,143.24) --
	(387.71,134.81) --
	(390.55,133.80) --
	(393.39,135.75) --
	(396.23,137.51) --
	(399.07,135.80) --
	(401.91,130.45) --
	(404.75,132.80) --
	(407.60,136.90) --
	(410.44,137.73) --
	(413.28,138.35) --
	(416.12,138.71) --
	(418.96,134.35) --
	(421.80,137.02) --
	(424.64,137.94) --
	(427.48,134.70);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt ,line join=round,line cap=round] ( 72.43,137.81) --
	( 75.27,136.65) --
	( 78.11,143.49) --
	( 80.95,137.43) --
	( 83.79,130.64) --
	( 86.63,140.10) --
	( 89.47,133.95) --
	( 92.31,138.95) --
	( 95.15,138.13) --
	( 97.99,134.38) --
	(100.83,138.70) --
	(103.67,136.65) --
	(106.51,142.33) --
	(109.35,140.08) --
	(112.19,140.76) --
	(115.03,135.30) --
	(117.87,131.53) --
	(120.71,134.31) --
	(123.55,136.31) --
	(126.40,138.07) --
	(129.24,138.68) --
	(132.08,133.35) --
	(134.92,135.06) --
	(137.76,137.42) --
	(140.60,132.08) --
	(143.44,134.84) --
	(146.28,141.15) --
	(149.12,136.50) --
	(151.96,137.75) --
	(154.80,136.26) --
	(157.64,132.36) --
	(160.48,135.15) --
	(163.32,133.44) --
	(166.16,135.02) --
	(169.00,137.46) --
	(171.84,137.44) --
	(174.68,138.85) --
	(177.52,137.64) --
	(180.36,136.64) --
	(183.20,136.98) --
	(186.04,135.19) --
	(188.88,135.00) --
	(191.72,137.46) --
	(194.56,133.62) --
	(197.41,136.58) --
	(200.25,133.35) --
	(203.09,137.85) --
	(205.93,138.76) --
	(208.77,132.56) --
	(211.61,136.25) --
	(214.45,135.48) --
	(217.29,136.00) --
	(220.13,134.15) --
	(222.97,138.92) --
	(225.81,140.34) --
	(228.65,137.72) --
	(231.49,138.74) --
	(234.33,140.75) --
	(237.17,137.08) --
	(240.01,139.13) --
	(242.85,132.23) --
	(245.69,135.17) --
	(248.53,135.31) --
	(251.37,134.56) --
	(254.21,137.20) --
	(257.05,135.96) --
	(259.89,133.73) --
	(262.73,140.13) --
	(265.57,133.33) --
	(268.42,134.06) --
	(271.26,136.71) --
	(274.10,138.64) --
	(276.94,143.32) --
	(279.78,136.35) --
	(282.62,136.03) --
	(285.46,138.94) --
	(288.30,139.04) --
	(291.14,138.48) --
	(293.98,133.60) --
	(296.82,137.06) --
	(299.66,135.32) --
	(302.50,136.59) --
	(305.34,139.19) --
	(308.18,132.33) --
	(311.02,130.50) --
	(313.86,133.96) --
	(316.70,137.48) --
	(319.54,136.37) --
	(322.38,135.54) --
	(325.22,141.95) --
	(328.06,137.28) --
	(330.90,136.24) --
	(333.74,135.82) --
	(336.59,129.34) --
	(339.43,131.63) --
	(342.27,137.63) --
	(345.11,139.18) --
	(347.95,141.59) --
	(350.79,135.78) --
	(353.63,132.71) --
	(356.47,132.87) --
	(359.31,137.75) --
	(362.15,135.79) --
	(364.99,140.06) --
	(367.83,133.34) --
	(370.67,138.30) --
	(373.51,137.87) --
	(376.35,138.13) --
	(379.19,134.63) --
	(382.03,136.24) --
	(384.87,143.24) --
	(387.71,134.81) --
	(390.55,133.80) --
	(393.39,135.75) --
	(396.23,137.51) --
	(399.07,135.80) --
	(401.91,130.45) --
	(404.75,132.80) --
	(407.60,136.90) --
	(410.44,137.73) --
	(413.28,138.35) --
	(416.12,138.71) --
	(418.96,134.35) --
	(421.80,137.02) --
	(424.64,137.94) --
	(427.48,198.96);
\definecolor{drawColor}{RGB}{0,191,255}

\path[draw=drawColor,line width= 0.8pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (424.64, 55.15) -- (424.64,233.80);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,189.66) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,212.78) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,176.65) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,201.49) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,198.71) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,181.04) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,229.00) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,206.56) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,197.52) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,206.18) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,187.44) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,197.65) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,209.44) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,197.73) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,187.08) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,203.49) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,166.81) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,196.41) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,183.27) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,165.90) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,179.34) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,169.71) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,233.80) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,206.03) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,201.56) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,163.02) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,207.09) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,188.70) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,183.72) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,210.89) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,183.23) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,176.55) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,179.09) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,193.27) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,178.26) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,203.34) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,215.62) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,193.21) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,158.42) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,209.95) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (441.68,134.70) -- (441.68,198.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (439.15,139.08) --
	(441.68,134.70) --
	(444.21,139.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (444.21,194.58) --
	(441.68,198.96) --
	(439.15,194.58);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.16,112.36) -- (433.16,164.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (430.63,116.74) --
	(433.16,112.36) --
	(435.69,116.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (435.69,159.81) --
	(433.16,164.19) --
	(430.63,159.81);
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (492.81,164.33) {$\displaystyle \hat{\alpha} = \frac{1}{n}\sum_{i=1}^{n} \hat{\alpha}_i= \;$5.22};
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (470.08,122.17) {$\alpha = 4.21$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (492.81, 77.15) {shock time point $T^*_1 + 1$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (364.99,161.69) {$y_{T_1^*+1}$};
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (335.16,196.46) {$\hat{y}^{2}_{T_1^*+1}=\hat{y}^{1}_{T_1^*+1} + \hat{\alpha}$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (379.19,198.96) -- (427.48,198.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.10,196.43) --
	(427.48,198.96) --
	(423.10,201.49);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (379.19,164.19) -- (427.48,164.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.10,161.66) --
	(427.48,164.19) --
	(423.10,166.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.75,198.96) rectangle (212.08,162.96);
\definecolor{drawColor}{RGB}{255,106,106}

\path[draw=drawColor,line width= 0.8pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] ( 75.75,186.96) -- ( 93.75,186.96);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.8pt,dash pattern=on 1pt off 3pt ,line join=round,line cap=round] ( 75.75,174.96) -- ( 93.75,174.96);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (102.75,183.51) {$\hat{y}_{t}^{1}$ without shock effects};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (102.75,171.51) {$\hat{y}_{t}^{2}$ with shock effects};
\end{scope}
\end{tikzpicture}

\end{document}