### 2020-08-09 manuscript/alphahist.tex
addition of simulation results

% Created by tikzDevice version 0.12.3.1 on 2020-08-08 20:35:31
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,252.94);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (301.08,  2.40) {$\hat{\alpha}_i$};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (  9.60,144.47) {Frequency};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.75, 48.00) -- (535.41, 48.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.75, 48.00) -- ( 66.75, 42.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (133.70, 48.00) -- (133.70, 42.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (200.65, 48.00) -- (200.65, 42.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (267.60, 48.00) -- (267.60, 42.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.56, 48.00) -- (334.56, 42.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (401.51, 48.00) -- (401.51, 42.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (468.46, 48.00) -- (468.46, 42.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (535.41, 48.00) -- (535.41, 42.00);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 66.75, 26.40) {7};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (133.70, 26.40) {8};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (200.65, 26.40) {9};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (267.60, 26.40) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (334.56, 26.40) {11};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (401.51, 26.40) {12};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (468.46, 26.40) {13};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (535.41, 26.40) {14};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 55.15) -- ( 48.00,233.80);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 55.15) -- ( 42.00, 55.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 80.67) -- ( 42.00, 80.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,106.19) -- ( 42.00,106.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,131.71) -- ( 42.00,131.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,157.23) -- ( 42.00,157.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,182.76) -- ( 42.00,182.76);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,208.28) -- ( 42.00,208.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,233.80) -- ( 42.00,233.80);

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 33.60, 55.15) {0};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 33.60, 80.67) {1};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 33.60,106.19) {2};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 33.60,131.71) {3};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 33.60,157.23) {4};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 33.60,182.76) {5};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 33.60,208.28) {6};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 33.60,233.80) {7};
\end{scope}
\begin{scope}
\path[clip] ( 48.00, 48.00) rectangle (554.16,240.94);
\definecolor{drawColor}{RGB}{0,191,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (109.11, 55.15) -- (133.70, 79.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (133.70, 79.74) -- (167.18,113.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (200.65,146.69) -- (234.13,180.17);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (211.31, 55.15) -- (234.13, 77.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.13, 77.96) -- (267.60,111.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (267.60,111.44) -- (287.88,131.71);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (301.08,144.91) -- (334.56,178.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (313.52, 55.15) -- (334.56, 76.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.56,178.39) -- (364.44,208.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.56, 76.19) -- (368.03,109.66);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (368.03,109.66) -- (401.51,143.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (415.72, 55.15) -- (434.98, 74.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (517.93, 55.15) -- (535.41, 72.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.75, 55.15) --
	(100.22, 55.15) --
	(100.22, 80.67) --
	( 66.75, 80.67) --
	( 66.75, 55.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (100.22, 55.15) --
	(133.70, 55.15) --
	(133.70, 80.67) --
	(100.22, 80.67) --
	(100.22, 55.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (133.70, 55.15) --
	(167.18, 55.15) --
	(167.18,131.71) --
	(133.70,131.71) --
	(133.70, 55.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (167.18, 55.15) --
	(200.65, 55.15) --
	(200.65,106.19) --
	(167.18,106.19) --
	(167.18, 55.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (200.65, 55.15) --
	(234.13, 55.15) --
	(234.13,233.80) --
	(200.65,233.80) --
	(200.65, 55.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (234.13, 55.15) --
	(267.60, 55.15) --
	(267.60,157.23) --
	(234.13,157.23) --
	(234.13, 55.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (267.60, 55.15) --
	(301.08, 55.15) --
	(301.08,131.71) --
	(267.60,131.71) --
	(267.60, 55.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (301.08, 55.15) --
	(334.56, 55.15) --
	(334.56,208.28) --
	(301.08,208.28) --
	(301.08, 55.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.56, 55.15) --
	(368.03, 55.15) --
	(368.03,208.28) --
	(334.56,208.28) --
	(334.56, 55.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (368.03, 55.15) --
	(401.51, 55.15) --
	(401.51,157.23) --
	(368.03,157.23) --
	(368.03, 55.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (401.51, 55.15) --
	(434.98, 55.15) --
	(434.98, 80.67) --
	(401.51, 80.67) --
	(401.51, 55.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (434.98, 55.15) --
	(468.46, 55.15) --
	(468.46, 55.15) --
	(434.98, 55.15) --
	(434.98, 55.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (468.46, 55.15) --
	(501.94, 55.15) --
	(501.94, 80.67) --
	(468.46, 80.67) --
	(468.46, 55.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (501.94, 55.15) --
	(535.41, 55.15) --
	(535.41, 80.67) --
	(501.94, 80.67) --
	(501.94, 55.15);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (282.53, 48.00) -- (282.53,240.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (322.70,226.14) -- (285.88,226.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (290.26,228.67) --
	(285.88,226.14) --
	(290.26,223.61);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (374.73,223.64) {$\displaystyle \hat{\alpha} = \frac{1}{n}\sum_{i=1}^{n} \hat{\alpha}_i= \;$10.22};
\end{scope}
\end{tikzpicture}

\end{document}