### 2020-08-09 manuscript/comp.tex

% Created by tikzDevice version 0.12.3.1 on 2020-08-08 20:33:27
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (578.16,252.94);
\begin{scope}
\path[clip] ( 48.00, 48.00) rectangle (554.16,240.94);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 72.43,119.70) --
	( 75.27,124.71) --
	( 78.11,116.82) --
	( 80.95,103.64) --
	( 83.79,121.66) --
	( 86.63,109.38) --
	( 89.47,132.22) --
	( 92.31,119.97) --
	( 95.15,121.43) --
	( 97.99,125.83) --
	(100.83,110.73) --
	(103.67,130.32) --
	(106.51,134.04) --
	(109.35,126.21) --
	(112.19,120.71) --
	(115.03,108.51) --
	(117.87,101.94) --
	(120.71,117.64) --
	(123.55,114.85) --
	(126.40,118.83) --
	(129.24,125.24) --
	(132.08,100.04) --
	(134.92,118.06) --
	(137.76, 98.96) --
	(140.60,107.61) --
	(143.44,117.58) --
	(146.28,121.77) --
	(149.12,111.50) --
	(151.96,118.34) --
	(154.80,108.34) --
	(157.64, 96.87) --
	(160.48,104.75) --
	(163.32,125.56) --
	(166.16,125.60) --
	(169.00,118.11) --
	(171.84,131.67) --
	(174.68,119.66) --
	(177.52,113.38) --
	(180.36,111.62) --
	(183.20,110.39) --
	(186.04,120.28) --
	(188.88,121.72) --
	(191.72,110.88) --
	(194.56,104.55) --
	(197.41,102.32) --
	(200.25,103.20) --
	(203.09,118.38) --
	(205.93,116.19) --
	(208.77,107.70) --
	(211.61,125.73) --
	(214.45,113.93) --
	(217.29,116.50) --
	(220.13,121.22) --
	(222.97,122.01) --
	(225.81,116.15) --
	(228.65,123.94) --
	(231.49,134.95) --
	(234.33,117.09) --
	(237.17,134.93) --
	(240.01,108.75) --
	(242.85,107.21) --
	(245.69,114.25) --
	(248.53,114.47) --
	(251.37,132.28) --
	(254.21,105.40) --
	(257.05,101.28) --
	(259.89,124.08) --
	(262.73,117.25) --
	(265.57,105.89) --
	(268.42,118.56) --
	(271.26,116.26) --
	(274.10,124.57) --
	(276.94,115.15) --
	(279.78,114.21) --
	(282.62,117.93) --
	(285.46,120.53) --
	(288.30,120.78) --
	(291.14, 93.84) --
	(293.98,104.20) --
	(296.82,121.77) --
	(299.66,120.74) --
	(302.50,125.57) --
	(305.34, 98.65) --
	(308.18,118.24) --
	(311.02,113.92) --
	(313.86,111.47) --
	(316.70,114.87) --
	(319.54,111.66) --
	(322.38,122.13) --
	(325.22,120.92) --
	(328.06,119.13) --
	(330.90,107.34) --
	(333.74,110.31) --
	(336.59,102.29) --
	(339.43,122.78) --
	(342.27,118.97) --
	(345.11,127.18) --
	(347.95,115.75) --
	(350.79, 98.94) --
	(353.63,103.64) --
	(356.47,119.92) --
	(359.31,111.38) --
	(362.15,117.39) --
	(364.99, 99.61) --
	(367.83,127.65) --
	(370.67,117.72) --
	(373.51,116.14) --
	(376.35,118.69) --
	(379.19,125.02) --
	(382.03,126.74) --
	(384.87,128.03) --
	(387.71,108.51) --
	(390.55,116.45) --
	(393.39,116.57) --
	(396.23,110.71) --
	(399.07,105.66) --
	(401.91,101.83) --
	(404.75,112.58) --
	(407.60,119.09) --
	(410.44,116.04) --
	(413.28,121.05) --
	(416.12,111.62) --
	(418.96,110.49) --
	(421.80,110.21) --
	(424.64,109.63) --
	(427.48,182.07);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.75, 48.00) -- (492.81, 48.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.75, 48.00) -- ( 66.75, 42.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (208.77, 48.00) -- (208.77, 42.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (350.79, 48.00) -- (350.79, 42.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (492.81, 48.00) -- (492.81, 42.00);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 66.75, 26.40) {0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (208.77, 26.40) {50};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (350.79, 26.40) {100};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (492.81, 26.40) {150};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 64.28) -- ( 48.00,201.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 64.28) -- ( 42.00, 64.28);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,110.08) -- ( 42.00,110.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,155.88) -- ( 42.00,155.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00,201.67) -- ( 42.00,201.67);

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 33.60, 64.28) {-5};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 33.60,110.08) {0};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 33.60,155.88) {5};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 33.60,201.67) {10};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 48.00, 48.00) --
	(554.16, 48.00) --
	(554.16,240.94) --
	( 48.00,240.94) --
	( 48.00, 48.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (578.16,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (301.08,  2.40) {Time $T$};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (  9.60,144.47) {Response};
\end{scope}
\begin{scope}
\path[clip] ( 48.00, 48.00) rectangle (554.16,240.94);
\definecolor{drawColor}{RGB}{255,106,106}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] ( 72.43,116.65) --
	( 75.27,115.78) --
	( 78.11,120.87) --
	( 80.95,116.36) --
	( 83.79,111.31) --
	( 86.63,118.35) --
	( 89.47,113.78) --
	( 92.31,117.49) --
	( 95.15,116.88) --
	( 97.99,114.09) --
	(100.83,117.31) --
	(103.67,115.79) --
	(106.51,120.01) --
	(109.35,118.34) --
	(112.19,118.84) --
	(115.03,114.78) --
	(117.87,111.97) --
	(120.71,114.04) --
	(123.55,115.53) --
	(126.40,116.84) --
	(129.24,117.29) --
	(132.08,113.33) --
	(134.92,114.60) --
	(137.76,116.36) --
	(140.60,112.38) --
	(143.44,114.43) --
	(146.28,119.13) --
	(149.12,115.67) --
	(151.96,116.60) --
	(154.80,115.49) --
	(157.64,112.59) --
	(160.48,114.67) --
	(163.32,113.39) --
	(166.16,114.57) --
	(169.00,116.38) --
	(171.84,116.37) --
	(174.68,117.42) --
	(177.52,116.52) --
	(180.36,115.78) --
	(183.20,116.03) --
	(186.04,114.70) --
	(188.88,114.55) --
	(191.72,116.39) --
	(194.56,113.53) --
	(197.41,115.73) --
	(200.25,113.33) --
	(203.09,116.67) --
	(205.93,117.35) --
	(208.77,112.74) --
	(211.61,115.48) --
	(214.45,114.91) --
	(217.29,115.30) --
	(220.13,113.92) --
	(222.97,117.48) --
	(225.81,118.53) --
	(228.65,116.58) --
	(231.49,117.34) --
	(234.33,118.84) --
	(237.17,116.11) --
	(240.01,117.63) --
	(242.85,112.50) --
	(245.69,114.69) --
	(248.53,114.79) --
	(251.37,114.23) --
	(254.21,116.19) --
	(257.05,115.27) --
	(259.89,113.61) --
	(262.73,118.37) --
	(265.57,113.31) --
	(268.42,113.85) --
	(271.26,115.83) --
	(274.10,117.26) --
	(276.94,120.74) --
	(279.78,115.56) --
	(282.62,115.32) --
	(285.46,117.49) --
	(288.30,117.56) --
	(291.14,117.14) --
	(293.98,113.51) --
	(296.82,116.09) --
	(299.66,114.80) --
	(302.50,115.74) --
	(305.34,117.67) --
	(308.18,112.57) --
	(311.02,111.21) --
	(313.86,113.78) --
	(316.70,116.40) --
	(319.54,115.57) --
	(322.38,114.95) --
	(325.22,119.73) --
	(328.06,116.25) --
	(330.90,115.48) --
	(333.74,115.16) --
	(336.59,110.34) --
	(339.43,112.05) --
	(342.27,116.51) --
	(345.11,117.67) --
	(347.95,119.46) --
	(350.79,115.14) --
	(353.63,112.85) --
	(356.47,112.97) --
	(359.31,116.60) --
	(362.15,115.14) --
	(364.99,118.32) --
	(367.83,113.32) --
	(370.67,117.01) --
	(373.51,116.69) --
	(376.35,116.89) --
	(379.19,114.28) --
	(382.03,115.47) --
	(384.87,120.69) --
	(387.71,114.41) --
	(390.55,113.66) --
	(393.39,115.12) --
	(396.23,116.42) --
	(399.07,115.15) --
	(401.91,111.17) --
	(404.75,112.92) --
	(407.60,115.97) --
	(410.44,116.59) --
	(413.28,117.05) --
	(416.12,117.32) --
	(418.96,114.07) --
	(421.80,116.06) --
	(424.64,116.75) --
	(427.48,114.33);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 1pt off 3pt ,line join=round,line cap=round] ( 72.43,116.65) --
	( 75.27,115.78) --
	( 78.11,120.87) --
	( 80.95,116.36) --
	( 83.79,111.31) --
	( 86.63,118.35) --
	( 89.47,113.78) --
	( 92.31,117.49) --
	( 95.15,116.88) --
	( 97.99,114.09) --
	(100.83,117.31) --
	(103.67,115.79) --
	(106.51,120.01) --
	(109.35,118.34) --
	(112.19,118.84) --
	(115.03,114.78) --
	(117.87,111.97) --
	(120.71,114.04) --
	(123.55,115.53) --
	(126.40,116.84) --
	(129.24,117.29) --
	(132.08,113.33) --
	(134.92,114.60) --
	(137.76,116.36) --
	(140.60,112.38) --
	(143.44,114.43) --
	(146.28,119.13) --
	(149.12,115.67) --
	(151.96,116.60) --
	(154.80,115.49) --
	(157.64,112.59) --
	(160.48,114.67) --
	(163.32,113.39) --
	(166.16,114.57) --
	(169.00,116.38) --
	(171.84,116.37) --
	(174.68,117.42) --
	(177.52,116.52) --
	(180.36,115.78) --
	(183.20,116.03) --
	(186.04,114.70) --
	(188.88,114.55) --
	(191.72,116.39) --
	(194.56,113.53) --
	(197.41,115.73) --
	(200.25,113.33) --
	(203.09,116.67) --
	(205.93,117.35) --
	(208.77,112.74) --
	(211.61,115.48) --
	(214.45,114.91) --
	(217.29,115.30) --
	(220.13,113.92) --
	(222.97,117.48) --
	(225.81,118.53) --
	(228.65,116.58) --
	(231.49,117.34) --
	(234.33,118.84) --
	(237.17,116.11) --
	(240.01,117.63) --
	(242.85,112.50) --
	(245.69,114.69) --
	(248.53,114.79) --
	(251.37,114.23) --
	(254.21,116.19) --
	(257.05,115.27) --
	(259.89,113.61) --
	(262.73,118.37) --
	(265.57,113.31) --
	(268.42,113.85) --
	(271.26,115.83) --
	(274.10,117.26) --
	(276.94,120.74) --
	(279.78,115.56) --
	(282.62,115.32) --
	(285.46,117.49) --
	(288.30,117.56) --
	(291.14,117.14) --
	(293.98,113.51) --
	(296.82,116.09) --
	(299.66,114.80) --
	(302.50,115.74) --
	(305.34,117.67) --
	(308.18,112.57) --
	(311.02,111.21) --
	(313.86,113.78) --
	(316.70,116.40) --
	(319.54,115.57) --
	(322.38,114.95) --
	(325.22,119.73) --
	(328.06,116.25) --
	(330.90,115.48) --
	(333.74,115.16) --
	(336.59,110.34) --
	(339.43,112.05) --
	(342.27,116.51) --
	(345.11,117.67) --
	(347.95,119.46) --
	(350.79,115.14) --
	(353.63,112.85) --
	(356.47,112.97) --
	(359.31,116.60) --
	(362.15,115.14) --
	(364.99,118.32) --
	(367.83,113.32) --
	(370.67,117.01) --
	(373.51,116.69) --
	(376.35,116.89) --
	(379.19,114.28) --
	(382.03,115.47) --
	(384.87,120.69) --
	(387.71,114.41) --
	(390.55,113.66) --
	(393.39,115.12) --
	(396.23,116.42) --
	(399.07,115.15) --
	(401.91,111.17) --
	(404.75,112.92) --
	(407.60,115.97) --
	(410.44,116.59) --
	(413.28,117.05) --
	(416.12,117.32) --
	(418.96,114.07) --
	(421.80,116.06) --
	(424.64,116.75) --
	(427.48,207.97);
\definecolor{drawColor}{RGB}{0,191,255}

\path[draw=drawColor,line width= 0.8pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (424.64, 55.15) -- (424.64,233.80);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,201.04) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,218.18) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,191.25) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,209.83) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,207.72) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,194.89) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,230.23) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,213.99) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,206.98) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,213.37) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,199.50) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,207.00) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,215.65) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,207.03) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,198.95) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,211.22) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,184.07) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,206.14) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,196.32) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,183.41) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,193.49) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,186.22) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,233.80) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,213.16) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,209.89) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,181.26) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,214.08) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,200.26) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,196.75) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,216.85) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,196.21) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,191.17) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,193.23) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,203.75) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,192.70) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,211.24) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,220.14) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,203.80) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,178.03) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (421.80,215.83) circle (  2.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (436.00,114.33) -- (436.00,207.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.47,118.72) --
	(436.00,114.33) --
	(438.53,118.72);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (438.53,203.59) --
	(436.00,207.97) --
	(433.47,203.59);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (487.13,158.65) {$\displaystyle \hat{\alpha} = \frac{1}{n}\sum_{i=1}^{n} \hat{\alpha}_i= \;$10.22};
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (464.40,114.33) -- (427.48,114.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (431.86,116.86) --
	(427.48,114.33) --
	(431.86,111.81);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (498.49,111.84) {$\hat{y}^{1}_{T_1^*+1} = 0.46$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (492.81, 70.86) {shock time point $T^*_1 + 1$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (364.99,179.57) {$y_{T_1^*+1}$};
\definecolor{drawColor}{RGB}{255,0,255}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (335.16,205.47) {$\hat{y}^{2}_{T_1^*+1}=\hat{y}^{1}_{T_1^*+1} + \hat{\alpha}$};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (379.19,207.97) -- (427.48,207.97);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.10,205.44) --
	(427.48,207.97) --
	(423.10,210.50);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (379.19,182.07) -- (427.48,182.07);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (423.10,179.54) --
	(427.48,182.07) --
	(423.10,184.60);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 66.75,207.97) rectangle (212.08,171.97);
\definecolor{drawColor}{RGB}{255,106,106}

\path[draw=drawColor,line width= 0.8pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] ( 75.75,195.97) -- ( 93.75,195.97);
\definecolor{drawColor}{RGB}{255,0,255}

\path[draw=drawColor,line width= 0.8pt,dash pattern=on 1pt off 3pt ,line join=round,line cap=round] ( 75.75,183.97) -- ( 93.75,183.97);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (102.75,192.53) {$\hat{y}_{t}^{1}$ without shock effects};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (102.75,180.53) {$\hat{y}_{t}^{2}$ with shock effects};
\end{scope}
\end{tikzpicture}

\end{document}