### 2020-12-10 Revision/list.tex
revision comments for the first reviewer added

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amscd}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{graphicx,times}
\usepackage{mathtools}
\usepackage{url}
%\usepackage{setspacing}
\usepackage{fullpage}
\usepackage{palatino}
\usepackage{mathpazo}
\usepackage{enumitem}
\usepackage{tcolorbox}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{xifthen} % Allows us to put in optional arguments to questions
\usepackage{color}
\usepackage{manfnt}
\usepackage{ifthen}
\usepackage{listings}
\usepackage{nicefrac,mathtools}
\usepackage[top=2.2cm,bottom=2.2cm,right=2.1cm,left=2.1cm]{geometry}
\usepackage{color}   %May be necessary if you want to color links
\usepackage{hyperref}
\usepackage{float}
\usepackage{showexpl}
\usepackage{xcolor}
\usepackage{mathtools}
\usepackage{caption}
\usepackage{chngcntr}
\usepackage{xparse}
\usepackage{etoolbox}
\usepackage{blkarray}
\usepackage{cprotect}

\DeclarePairedDelimiter\abs{\lvert}{\rvert}
\DeclarePairedDelimiter\norm{\lVert}{\rVert}

\newcommand{\R}{\mathbb{R}}
\newcommand{\ds}{\displaystyle}
\newcommand{\mustar}{\mu^{\textstyle{*}}}
\newcommand{\betastar}{\hat{\beta}^{\textstyle{*}}}
\newcommand{\betahat}{\hat{\beta}}
\newcommand{\betaw}{\hat{\beta}_w}
\newcommand{\betastarT}{\hat{\beta}^{\textstyle{*}^T}}
\newcommand{\betabar}{\bar{\beta}^{\textstyle{*}}}
\newcommand{\bstar}{b^{\textstyle{*}}}
\newcommand{\vstar}{v^{\textstyle{*}}}
\newcommand{\Bstar}{B^{\textstyle{*}}}
\newcommand{\wstar}{w^{\textstyle{*}}}
\newcommand{\BICstar}{\textsc{bic}^{\textstyle{*}}}
\newcommand{\minBICstar}{\min_{s=1,...,r}\left\{\BIC^{\textstyle{*}}(s)\right\}}
\newcommand{\sstar}{s_m^{\textstyle{*}}}
\newcommand{\epstar}{\varepsilon^{\textstyle{*}}}
\newcommand{\epstarT}{\varepsilon^{{\textstyle{*}^T}}}
\newcommand{\epresstar}{\widehat{\varepsilon}^{\textstyle{*}}}
\newcommand{\epresstarT}{\widehat{\varepsilon}^{\textstyle{*}^T}}
\newcommand{\residstar}{\widehat{\varepsilon}^{\textstyle{*}}}
\newcommand{\Ystar}{Y^{\textstyle{*}}}
\newcommand{\Xstar}{X^{\textstyle{*}}}
\newcommand{\Wstar}{W^{\textstyle{*}}}
\newcommand{\Wstarinv}{W^{{\textstyle{*}^{-1}}}}
\newcommand{\Zstar}{Z^{\textstyle{*}}}
\newcommand{\YstarT}{Y^{{\textstyle{*}^T}}}
\newcommand{\XstarT}{X^{\textstyle{*}^T}}
\newcommand{\Sigstar}{\widehat{\Sigma}^{\textstyle{*}}}
\newcommand{\Sigstarhalf}{\widehat{\Sigma}^{\textstyle{*}^{1/2}}}
\newcommand{\Sigstarhalfinv}{\widehat{\Sigma}^{\textstyle{*}^{-1/2}}}
\newcommand{\lstar}{l^{\textstyle{*}}}

\newcommand{\Astar}{A^{\textstyle{*}}}
\newcommand{\Gostar}{\widehat{G}_o^{\textstyle{*}}}
\newcommand{\GostarT}{\widehat{G}_o^{\textstyle{*}^T}}
\newcommand{\Gohat}{\widehat{G}_o}
\newcommand{\GohatT}{\widehat{G}_o^T}
\newcommand{\Sigresstar}{\widehat{\Sigma}^{\textstyle{*}}}

\newcommand{\B}{\mathcal{B}}
\newcommand{\Lnorm}{\mathcal{L}}
\newcommand{\Sub}{\mathcal{S}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\Proj}{\mathcal{P}}
\newcommand{\Env}{\mathcal{E}}
\newcommand{\Envspace}{\Env_{\Sigma}(\B)}
\newcommand{\utrue}{u_{\text{true}}}
\newcommand{\BIC}{\textsc{bic}}
\newcommand{\dimEnv}{\text{dim}\{\Envspace\}}
\newcommand{\minBIC}{\min_{s=1,...,r}\left(\BIC(s)\right)}
\newcommand{\nboot}{n_{\text{boot}}}

\newcommand{\X}{\mathbb{X}}
\newcommand{\Y}{\mathbb{Y}}
\newcommand{\Ymatstar}{\Y^{\textstyle{*}}}
\newcommand{\YmatstarT}{\Y^{\textstyle{*}^T}}
\newcommand{\Xmatstar}{\X^{\textstyle{*}}}
\newcommand{\XmatstarT}{\X^{\textstyle{*}^T}}

\newcommand{\Sigres}{\widehat{\Sigma}}
\newcommand{\SigY}{\widehat{\Sigma}_{Y}}
\newcommand{\SigX}{\widehat{\Sigma}_{X}}
\newcommand{\Pu}{\widehat{\Proj}_{\Env_u}}
\newcommand{\Pj}{\widehat{\Proj}_{\Env_j}}
\newcommand{\Qu}{\widehat{\Q}_{\Env_u}}
\newcommand{\Qj}{\widehat{\Q}_{\Env_j}}
\newcommand{\PD}{\widehat{\Proj}_{D}}
\newcommand{\betau}{\hat{\beta}_u}
\newcommand{\betaj}{\hat{\beta}_j}
\newcommand{\res}{\widehat{\varepsilon}}

\newcommand{\sestar}{\text{se}^{\textstyle{*}}\{\text{vec}(\hat{\beta})\}}
\newcommand{\setrue}{\text{se}_{\text{true}}\{\text{vec}(\hat{\beta})\}}

\newcommand{\vecop}[1]{\text{vec}\left( #1 \right)}
\newcommand{\vechop}[1]{\text{vech}\left( #1 \right)}
\newcommand{\indep}{\rotatebox[origin=c]{90}{$\models$}}

\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\Var}{var}

\newtheorem{lem}{Lemma} 
\newtheorem{thm}{Theorem} 
\allowdisplaybreaks

\setlength{\parindent}{0cm}

\newcommand{\response}[1]{\noindent \textcolor{blue}{\emph{Response:} #1}}
\cMakeRobust\response

\title{Referee report checklist}
\author{Jilei Lin and Daniel J. Eck}
\date{}

\bibliographystyle{plainnat}

\begin{document}

\maketitle

Thank you for the opportunity to resubmit at \emph{International Journal of Forecasting}. Thank you to the reviewers for comments that led to an improved version of this manuscript. We address all of the reviewers' comments below. An outlines of each change made or a rebuttal is provided for all of the reviewer's comments.

\newpage
\section*{Reviewer 1: \texttt{Cover letter for authors.pdf}}

Thank you very much for taking the time to provide a thoughtful review of our manuscript. We think that these comments will go a long way to improve the exposition of our manuscript. \\

%% addressed
{\bf  Comment 1:} How could the method be applied to forecast two or more period after a shock? \\

\response{The method can be extended to forecast two or more periods after a shock. There are two situations.  \\

First, suppose that there is no additional shock after $T_1^* + 1$. From equation (1) in the manuscript, the forecast at time point $T_1^*+2$ is
\begin{align*}
 \hat{y}_{i, T_i^* + 2} = \hat{\eta}_1 + \hat{\phi}_i \hat{y}_{1, T_1^*+1}^{2} + \hat{\theta}_i'\mathbf{x}_{1, T_1^*+2}, \tag{*}
\end{align*}
where $\hat{y}_{1, T_1^*+1}^{2}=\hat{y}_{1, T_1^*+1}^{1} + \hat{\alpha}_{T_1^* + 1}$, where $\hat{y}_{1, T_1^*+1}^{1}$ is the forecast without adjustment, and  $\hat{\alpha}_{T_1^* + 1}$ is the shock-effect estimator for the shock at $T_1^*+1$. The forecast for $d>2$ more periods is
\begin{align*}
 \hat{y}_{i, T_i^* + d} = \hat{\eta}_1 + \hat{\phi}_i \hat{y}_{1, T_1^*+d-1}+ \hat{\theta}_i'\mathbf{x}_{1, T_1^*+d},
\end{align*}
where $\hat{y}_{1, T_1^*+d-1}$ can be also computed recursively using the above formula and we  emphasize  $ \hat{y}_{i, T_i^* + 1}=\hat{y}_{1, T_1^*+1}^{2}$. Since we use AR(1) model, $|\phi_1 |< 1$ is assumed.  As one can expect, the impact of the shock  $\alpha_1$ will diminish as $d$ increases. Similar logic can apply for AR($p$). \\

Second, suppose that  there is one more shock effect after $T_1^*+1$, say, $\alpha_{T_1^* + d^*}$ for some $d^* > 1$. Without loss of generality, assume $d^* = 2$. Under our framework, users may want to construct another set of donor pool with time series that experienced shock effects similar to $\alpha_{T_1^* + 2}$. Next,  apply our proposed method to compute $\hat{\alpha}_{T_1^* + 2}$, the estimator for the shock effect at $T_1^* +2$. The forecast at the point $T_1^* + 2$ will be,  
\begin{align*}
 \hat{y}_{i, T_i^* + 2} = \hat{\eta}_1 + \hat{\phi}_i \hat{y}_{1, T_1^*+1}^{2} + \hat{\theta}_i'\mathbf{x}_{1, T_1^*+2} + \hat{\alpha}_{T_1^* + 2}.
\end{align*}
These two cases can be further combined to generalize forecasting for two or more periods. \\
}

{\bf Comment 2:} Page 3, line 30: \citet{blundell1998initial} model is a fixed effect dynamic panel and estimate the parameters using the panel analysis. It would be nice if you clarify in what sense your work is similar to them. \\

\response{The model of \citet{blundell1998initial} is
\begin{align*}
  y_{it} = \alpha y_{i, t-1} + \beta_1'x_{it} + \beta_2'x_{it-1} + \eta_i + v_{it}, \quad i = 1, \ldots, N, \quad t = 2, \ldots, T.
\end{align*}
where $|\alpha|<1$. In contrast, our model is
\begin{align*}
 y_{i,t} =\eta_i +\alpha_i D_{i,t} + \phi_i y_{i, t-1} + \theta_i'\mbf{x}_{i,t} + \varepsilon_{i,t},
\end{align*}
where $|\phi_i| < 1$. Our model is similar to the one of \citet{blundell1998initial}  in the sense that we have the fixed-effect decomposition of the error term $\eta_i +\varepsilon_{i,t}$ and the autoregressive structure. However, our model allows for $\eta_i$ as random effects, different time lengths for each time series, different autoregressive parameters, and a shock component.}\\

{\bf Comment 3:} Page 4, line 50: Your paper considers forecasting a time-series model. It is confusing for readers to read your sentence ”In this article, we consider a dynamic panel data model...''.\\

\response{Sorry for the confusion. We will correct that in the revised version.}\\

{\bf  Comment 4:} Figures on page 5: Figure labels, (a) and (b), are missing. What does the black line in the top graph represent? Is it the realized observations?\\

\response{Thanks for pointing it out. The labels (a) and (b) are now added in the revised manuscript.   The \textcolor{black}{black} line stands for the realized observations whereas the \textcolor{magenta}{magenta} line stands for the fitted values.} \\

{\bf Comment 5:} Page 7, line 52: Are these time series estimates or panel data estimates? If panel estimators are used, then OLS estimators in this model results in inconsistency of the estimators. In that case, how wouldn't this inconsistency affect the results in section 3 about the unbiasedness? \\

\response{Sorry for not making it clear. They are time series estimates or the ordinary least squares (OLS) estimates but not panel data estimates. In fact, our method can be generalized to any model whose parameter can be estimated in an unbiased fashion. In the second paragraph of Section 6, we wrote \\

\begin{tcolorbox}
``Although our work is developed for time-series or AR($p$) models, in fact, it can be generalized to any similar setting with a model of the response, whose parameters can be estimated \emph{\textcolor{red}{unbiasedly}}, an additive shock-effect structure, and the structure that the time series in the donor pool are independent of the one of interest.''
\end{tcolorbox}

\vspace{0.4cm}In other words, if the estimator used in panel data is unbiased, our method can adapt to this scenario by weighting the shock effects estimated by panel data estimation procedure. In this case, the propositions in Section 3 still hold though the exact variance expressions will change since we are not using OLS estimators.}\\

{\bf Comment 6:} Page 8, line 46: It is not clear to met how you estimate $\alpha_i$'s by OLS. Could you give the expressions for $\hat{\alpha}_i$? Also, how can you identify $\alpha_i$ from $\eta_i$? Isn't $\hat{\alpha}_{\rm adj}$ an estimator of $E(\alpha_1)$? \\

\response{Sorry for not making it clear. The estimation procedures can be explained in detail as follows. Note that for $i = 2, \ldots, n+1$, the design matrix for $i$th time series is $(\mathbf{1}_{T_i}, \mathbf{D}_{i}, \mathbf{X}_{i})$, where $\mathbf{X}_i$ is the covariates  of $i$th time series and $\mathbf{D}_{i}=(D_{i,t})_{t=1}^{T_i}$ with $D_{i,t}=I(t=T_i^*+1)$. We define $\mathbf{U}_i = (\mathbf{1}_{T_i}, \mathbf{D}_{i}, \mathbf{X}_{i})$. Then the expression for $\hat{\alpha}_i$ is
\begin{align*}
  \hat{\alpha}_i = [(\mathbf{U}'_i\mathbf{U}_i)^{-1}\mathbf{U}_i'\mathbf{Y}_{i}]_{2}, \quad 
  \text{ where }\quad \mathbf{Y}_{i}=(y_{i,t})_{t=1}^{T_i} 
 \quad  \text{ for } i = 2, \ldots, n+1.
\end{align*}
Note that $[\cdot]_{2}$ denotes the second element of a vector. That is, $\hat{\alpha}_i$ is the second element of the OLS estimate vector. Recall that our model is
\begin{align*}
  y_{i,t} = \eta_i + \alpha_i D_{i,t} + \phi_i y_{i,t-1} + \theta_i'\mathbf{x}_{i,t} + \varepsilon_{i,t}, 
  \quad i = 1, \ldots, n+1.
\end{align*}
$\eta_i$ represents the effect of the intercept. Thus, we cannot identify $\alpha_i$ from $\eta_i$. As defined in (4) in the manuscript, through averaging, $\hat{\alpha}_{\rm adj}$ is an estimator  of $E(\alpha_1)$.} \\

{\bf Comment 7:} Page 9, line 10: It is not clear to me what $\mathbf{U}_i$'s are. An example could help understanding that better. \\

\response{Thanks for pointing it out. We will clarify it in the revised version. $\mathbf{U}_i$'s represent the design matrix for the OLS estimation at the $i$th time series. For example, suppose the sample size for $i$th time series in the donor pool (i.e., $i\geq 2$) is 4; we have one covariate consisting of $x_1, \ldots, x_4$; and the shock occurs at $t = 4$. In this case, $\mathbf{U}_i$ is 
\begin{align*}
  \mathbf{U}_i = \begin{pmatrix}
    1 & 0 & x_1 \\
    1 & 0 & x_2 \\
    1 & 0 & x_3 \\
    1 & 1 & x_4
  \end{pmatrix}.
\end{align*}
Suppose the sample size for the time series of interest is also 4, and we have a covariate $k_1, \ldots, k_4$. Note that the design matrix of the time series of interest will not contain $D_{1,t}$ since $D_{1,t}= I(t=T_1^*+1)=0$ due to we do not observe the shock. In this case, 
\begin{align*}
  \mathbf{U}_1 = \begin{pmatrix}
    1 &  k_1 \\
    1 &  k_2 \\
    1 &  k_3 \\
    1 &  k_4
  \end{pmatrix}.
\end{align*}} \\

{\bf Comment 8:} Page 9, line 15: Why the closed form expressions for $E(\hat{\alpha}_{\rm IVW})$ and $Var(\hat{\alpha}_{\rm IVW})$ are not provided? \\

\response{Recall that in page 8, we wrote \\
\begin{tcolorbox}
''The inverse-variance weighted estimator is defined as 
\begin{align*}
  \hat{\alpha}_{\rm IVW} = \frac{\sum_{i=2}^{n+1} \hat{\alpha}_i / \hat{\sigma}_{i\alpha}^2}{\sum_{i=2}^{n+1} 1/\hat{\sigma}_{i\alpha}^2},
  \quad \text{ where } \quad  \hat{\sigma}_{i\alpha}^2 = \hat{\sigma}^2_i( \mathbf{U}_i'\mathbf{U}_i)_{22}^{-1},
\end{align*}
where  $\hat{\alpha}_i$ is the OLS estimator of $\alpha_i$, 
$\hat{\sigma}_i$ is the residual standard error from OLS estimation, 
and $\mathbf{U}_i$ is the design matrix for OLS with respect to time series 
for $i = 2, \ldots, n+1$. Note that since $\sigma$ is unknown, estimation 
is required and the numerator and denominator terms are dependent in general.''
\end{tcolorbox} \\
To be more specific, $\hat{\sigma}_{i\alpha}$ depends on $\hat{\alpha}_i$. As a result, the numerator and denominator of $ \hat{\alpha}_{\rm IVW}$ are dependent. Without making distributional assumptions, it is impossible to evaluate $E(\hat{\alpha}_{\rm IVW})$ and $Var(\hat{\alpha}_{\rm IVW})$ in a closed form.}\\

{\bf Comment 9:} Page 13, line 14: $\mu_{\alpha}$ is unknown, and in practice we have to estimate it. Would the results hold if one replaces $\mu_{\alpha}$ with its estimate? \\

\response{We go for a plug-in approach. If we can estimate $\mu_{\alpha}$ well, the results should hold with high probability. Section 3.2 is dedicated to this estimation problem. Using plug-in approaches, we get a decision rule to determine whether the scalar adjustment will be beneficial with correctness probability $p$. Section 3.3 details the leave-one-out cross validation procedures, which are tailored to our setup, to estimate $p$. As a result, users can be informed about the credibility of our methods} \\

{\bf Comment 10:} Page 14, line 7: Shouldn’t $E(\alpha_1)$ be $\mu_{\alpha}$? Previously $\mu_{\alpha}$ was used.\\

\response{In the line 7 at Page 14, $E(\alpha_1)$ is taken under $\mathcal{M}_{21}$ or $\mathcal{M}_{22}$. In contrast, $\mu_{\alpha}$ is used under $\mathcal{M}_{1}$. Recall that in Section 2.1,
\begin{align*}
  &\alpha_i \sim \mathcal{F}_{\alpha} 
 \quad  \text{ with }\quad  E_{\mathcal{F}_{\alpha}}(\alpha_i) =\mu_{\alpha}, Var_{\mathcal{F}_{\alpha}}(\alpha_i)=\sigma^2_{\alpha},\quad   \text{iid under } \mathcal{M}_1\\
 & \alpha_i = \mu_{\alpha} +\delta_i'\mathbf{x}_{i, T_i^*+1} + \tilde{\varepsilon}_i\quad  \text{iid under } \mathcal{M}_{21} \text{ or } \mathcal{M}_{22}.
\end{align*}
In other words,  $\mathcal{M}_{21}$  or $\mathcal{M}_{22}$ is an extended version of $\mathcal{M}_{1}$ by allowing it to depend on some covariates. However, in general, $E(\alpha_1)\neq \mu_{\alpha}$ under   $\mathcal{M}_{21}$  or $\mathcal{M}_{22}$ due to the presence of $\delta_i'\mathbf{x}_{i, T_i^*+1}$.} \\

{\bf Comment 11:} Page 18, numerical example: The model setup does not contain the individual effects (intercepts) similar to the models in equations (1)-(2) on page 6. How would the result change if you include them with different variances? \\

\response{The result wouldn't change. In fact, in our proof listed in the Appendix, we do not make use of the assumption that $\eta_i$ follows the same distribution. In other words, the results  still hold when the individual effects have different variances.}\\

{\bf Comment 12:} Page 23, construction of donor pool: Your model assumes that the shocks have the same distribution. Given that these shocks are from different sources and happened because of various reasons, how could one justify them? Practically, there are many previous shocks. How can someone choose from them?\\

\response{It is a good question. To clarify, $\mathcal{M}_1$ assumes that the shocks have the same distribution. However, $\mathcal{M}_{21}$ assumes the shocks have the same variance but have different conditional means with allowing covariates to play a role. Moreover, $\mathcal{M}_{22}$ assumes they have different conditional means and conditional variances by allowing $\delta_i$ to be random. But the shock effects are assumed to follow a general family of  unknown distribution with possible different means and variances.\\

$\mathcal{M}_{21}$ and $\mathcal{M}_{22}$ are constructed for generalization. In practice, it is hard to verify those assumptions. It leaves the discretion for users to construct donor pool by gathering time series experiencing similar events to the one the time series of interest experienced. \\

Theoretically, it is possible to verify the assumptions of $\mathcal{M}_{1}$ using empirical Bayes methods to some degree. First, we need a  pool of shock effects that users believe they are realizations from a common distribution, say, $\mathcal{F}_{\alpha}$. Using empirical Bayes methods, we can estimate $\mathcal{F}_{\alpha}$ to have $\hat{\mathcal{F}}_{\alpha}$. To verify whether the remaining shock effects are from $\mathcal{F}_{\alpha}$, we can check the quantile of $\hat{\mathcal{F}}_{\alpha}$ to judge whether they are extreme in $\hat{\mathcal{F}}_{\alpha}$. However, the theoretical justification for this approach is left for future research. \\

In practice, someone can select the shocks that are similar to the one of interest. For example, they can be similar in causes, scale, and characteristics. Using our data analysis example, on March 9th 2020, Conoco Phillips experienced a shock during COVID-19 pandemic and oil supply shock. During COVID-19 pandemic, the US economy is in a situation that resembles the one in financial crisis. As a result, we select the shocks in the past that are related to financial crisis and oil supply shock.} \\

{\bf Comment 13:} Figure 2 on Page 24: Are the black dots the real data? What does the line represent?\\

\response{Thanks for pointing it out. We will clarify this in the revised version. The black dots represent the real data. The line in \textcolor{magenta}{magenta} color represents the fitted line of our model.}

\bibliography{synthetic-prediction-notes}

\end{document}