### 2021-02-06 Revision/letter.tex
added letter

\documentclass[11pt]{article}

\usepackage{tikz}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{color}
\usepackage[export]{adjustbox}

\setlength{\parindent}{0em}
\setlength{\parskip}{1em}

\usepackage{geometry}
\geometry{margin=1in}

\renewcommand*\arraystretch{1.05}

\begin{document}

\footnotesize
%\color{lightblue}
\begin{tikzpicture}
  \node at (0,0) {\includegraphics[height=1.1in, keepaspectratio = true]{UI.png}};
  \node at (13.4, -0.05) {\vspace{1cm}\begin{tabular}{l}
      \\ \sc Jilei Lin \\
       \it Graduate Student \\
       \it Department of Statistics \\
       University of Illinois \\
       Daniels Hall, Room 0506A \\
	   1010 W. Green St \\
	   Urbana, IL 61801 \\
       \texttt{jileil2@illinois.edu} \\ 
       \end{tabular} };
\end{tikzpicture}

\normalsize

Dear editor,

We are grateful to submit a revision for our manuscript ``Minimizing post-shock forecasting error through aggregation of outside information'' at International Journal of Forecasting.  The responses to all of the comments are included in the resubmission files.

Thank you for your reconsideration.

Sincerely,

Jilei Lin \\
Daniel J. Eck

\end{document}